%%% License: Creative Commons Attribution Share Alike 4.0 (see https://creativecommons.org/licenses/by-sa/4.0/)
%%% Slides are based heavily on earlier versions of this course taught by Jesper Rudiger.

\documentclass[english,10pt
%,handout
,aspectratio=169
]{beamer}
\input{../FM_preamble.tex}



\title{Financial Markets Microstructure}

\author{Egor Starkov}

\date{K{\o}benhavns Unversitet \\
	Spring 2022}



\begin{document}
	
%\frame[plain]{\titlepage}
\addtocounter{framenumber}{-1}

\begin{frame}{Ch 7, ex 3}
	\textbf{Payments for order flow. }
	\begin{itemize}
		\item One risky security, pays either $v^{H}=\mu +\sigma $ or $v^{L}=\mu -\sigma $ with equal probabilities at $t=1$.
		\item At $t=0$ investor gives his broker an order to buy or sell one unit.
		\begin{itemize}
			\item W.p. $\phi$ \structure{retail investor} is retail investor, buys and sells randomly 50/50.
			\item W.p. $1-\phi$ \structure{institutional investor}, \alert{informed} w.p. $\alpha$. If uninformed, buys and sells randomly 50/50.
		\end{itemize}
		\item Three risk-neutral \structure{dealers} $1,2$, and $3$ post quotes before the broker contacts them.
		\begin{itemize}
			\item Broker cannot split his order among dealers
			\item Dealers have no private information on the payoff of the security
		\end{itemize}
	\end{itemize}
\end{frame}


\begin{frame}{Ch 7, ex 3, part a}
	\begin{exampleblock}{}
		(a) Assume that there is no payment for order flow between the
		broker and the three dealers. In this case, the broker randomly selects one
		dealer among those posting the best price for his order. Compute the bid and
		ask quotes posted by the dealers.
	\end{exampleblock}
	
	\pause
	
	If any single dealer posts quotes at a profit, the other two can undercut and attract order flow $\Rightarrow$ we are down to standard Glosten-Milgrom with competitive dealers.
	\begin{align*}
		a &= \mathbb{E}[v|\text{buy}] = \mu + \mathbb{P}(\text{informed}|\text{buy}) \sigma
		\\
		&= \mu + \frac{(1-\phi)\alpha/2}{(1-\phi)\alpha/2 + \left[\phi+(1-\phi)(1-\alpha) \right]/2} \sigma
		\\
		&= \mu + (1-\phi)\alpha \sigma
		\\
		b &= \mu - (1-\phi)\alpha \sigma
	\end{align*}
\end{frame}


\begin{frame}{Ch 7, ex 3, part b}
	\begin{exampleblock}{}
		(b) Assume now that dealer 1 has a payment for order flow
		arrangement under which the broker gives dealer 1 all orders from retail
		investors and the dealer commits to execute all these orders at the best
		quotes (i.e., the ask and bid price set by the remaining dealers, 2 and 3).
		Other orders are sent to dealer 2 or 3 as in question a. What are the quotes
		posted by dealers 2 and 3? Deduce that the bid-ask spread is higher in this
		case than where there is no payment for the order flow.
	\end{exampleblock}
	
	\pause
	
	Dealers 2 and 3 are still competing with each other $\Rightarrow$ post zero-profit quotes
	\begin{align*}
		a &= \mu + \mathbb{P}(\text{informed}|\text{buy,institutional}) \sigma
		\\
		&= \mu + \frac{(1-\phi)\alpha/2}{(1-\phi)\alpha/2 + (1-\phi)(1-\alpha)/2} \sigma
		\\
		&= \mu + \alpha \sigma > \mu + (1-\phi)\alpha \sigma
		\\
		b &= \mu - \alpha \sigma < \mu - (1-\phi)\alpha \sigma
	\end{align*}
\end{frame}


\begin{frame}{Ch 7, ex 3, part c}
	\begin{exampleblock}{}
		(c) Let $P$ be the payment of dealer 1 to the broker. What is the
		largest possible value of $P$?
	\end{exampleblock}
	
	\pause
	
	\begin{itemize}
		\item Dealer 1's per-order profit is:
		\begin{itemize}
			\item $a - \mu = \alpha \sigma$ from buy orders
			\item $\mu - b = \alpha \sigma$ from sell orders
		\end{itemize}
		\item so dealer 1 would be willing to pay up to $\alpha \sigma$ per order.
	\end{itemize}
\end{frame}


\begin{frame}{Ch 7, ex 3, part d}
	\begin{exampleblock}{}
		(d) Is payment for order flow beneficial or detrimental to investors?
	\end{exampleblock}
	
	\pause 
	
	\begin{itemize}
		\item All investors trade at worse prices under order flow payments -- bad.
		
		\item Dealer 1 and broker profit from order flow payments.
		
		\item If many brokers compete with each other, they may transmit part of the profit to (some?) investors.
	\end{itemize}
\end{frame}



\end{document} 